Demographic Factors and Climate Action: Understanding How Age and Education Shape Sustainable Choices

### 2024-12-15 paper/climate_change.qmd
slight adjustments for rendering

### Education
#### Individual 2018
@fig-educ shows the distribution of respondents by their highest level of education attained in 2018. The most common education levels are "Completed undergraduate degree" and "Postgraduate/professional school," indicating a well-educated respondent population. In contrast, fewer respondents reported having attended or completed community college, vocational, or trade school, suggesting lower representation in these categories. A small portion of respondents chose not to disclose their education level. Overall, the distribution reflects a skew toward higher levels of educational attainment.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-educ
#| fig-cap: "The histogram shows the distribution of education levels of 2018 survey respondents."

@@

```
#### Comparison 2018 to 2021
Table 3 illustrates the self-reported level of extent individuals feel informed about climate change and climate action for 2018 and 2021. In 2018, the large majority of individuals say they are very informed with 60 percent and the following category is not very informed with 27 percent. In 2021, The majority of individuals also say they are very informed but to a lesser degree with 48 percent and this is closely followed by only a little informed with 37 percent. There is a trend that individuals in 2018 reported greater levels of knowledge about climate change whereas 2021 illustrates a more even split between very informed and only little informed, portraying that fewer individuals have adequate knowledge about climate change.

@@

# Print the table (using kable for styling)
informed_summary_combined %>%
  kbl(caption = "This table highlights the extent feel informed distribution of survey respondents in 2018 and 2021") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

#### Comparison 2018 to 2021
fig-eight compares the preferred methods of delivering climate change information between the 2018 and 2021 surveys. While the top three methods namely advertising campaigns, newsletters, and the Toronto.ca website remained consistent across both years, there were notable shifts in preferences. Traditional methods such as newsletters, advertising campaigns, brochures, and pamphlets saw a decline in 2021. In contrast, social media platforms including Twitter, Facebook, and Instagram gained traction, reflecting a growing reliance on digital communication channels.

@@

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-eight
#| fig-cap: "This table highlights the prefered method of communication of survey respondents in 2018 and 2021."

@@

# Print the table (using kable for styling)
communication_summary_combined %>%
  kbl() %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

### Likelihood Taking Action 
fig-nine illustrates the self-reported likelihood of individuals taking various actions to minimize environmental impacts. The x-axis displays nine different climate actions, while the y-axis represents the percentage of individuals corresponding to each likelihood level. A stacked bar chart is used to visualize the distribution of responses, with a color-coded legend ranging from green (high likelihood of action) to red (low likelihood of action).

@@

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-nine
#| fig-cap: "The stacked bar plot shows the likelihood of survey respondents taking various actions to address climate change."

@@

#### Likelihood of switching to an electric or hybrid vehicle
fig-ten reveals that education is the primary predictor for the likelihood of switching to an electric or hybrid vehicle. Individuals with "Some Community/Trade School" or "Some University" education are "Very likely" to invest, though with moderate variability (error rate 63.8%). Those with higher education show more age-based segmentation. Younger individuals under 35 years are more likely to invest, with a lower error rate (30%). Older individuals, particularly those over 69 years, are "Very unlikely" to invest. This model demonstrates how age and education interact to influence vehicle choice.

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-ten
#| fig-cap: "Probability tree for likelihood of investing in an electric or hybrid vehicle, predicted by age and education."

@@

#### Likelihood of minimizing car use
fig-eleven shows that the likelihood of minimizing car use is consistent across the sample, with a high error rate (61.9%). The model classifies all individuals into the "Already doing this or have done this" category, indicating widespread adoption of this behavior, regardless of age or education

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-eleven
#| fig-cap: "Probability tree for likelihood of minimizing car use, predicted by age and education."

@@

#### Likelihood of walking or cycling shorter distances
fig-twelve indicates that education is the primary factor influencing short-distance walking or cycling. Those with higher education (postgraduate, undergraduate, or some university) are more likely to engage in this behavior, with an error rate of 51.9%. For those with lower education, age plays a significant role, with younger individuals (under 66 years) more likely to walk or cycle, with an error rate of 60%. This model highlights how education and age together shape short-distance mobility, with error rates reflecting the model's predictive variability.

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-twelve
#| fig-cap: "Probability tree for likelihood of walking or cycling shorter distances, predicted by age and education."

@@

### Waste and Product Actions
#### Likelihood of purchasing green products
fig-thirteen reveals that age is the initial split for green product purchasing behavior. Younger individuals (under 38 years) are more likely to purchase green products, with an error rate of 51.9%. Education further refines the prediction. Older individuals (over 38 years) are classified by both age and education, with postgraduate education leading to higher engagement in purchasing green products, with an error rate of 60%. This model shows how age and education influence consumer choices in sustainable products.
```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-thirteen
#| fig-cap: "Probability tree for likelihood of purchasing green products, predicted by age and education."

@@

#### Likelihood of reducing waste by repurposing used product
fig-fourteen shows that age is the primary factor for waste reduction behaviors. Younger individuals (under 35 years) are more likely to repurpose waste, with an error rate of 57.1%. Education further refines the classification. Older individuals (over 35 years) show varied likelihoods based on education, with more highly educated individuals more likely to repurpose waste, with an error rate of 60%. This highlights the influence of both age and education on waste-reduction habits.
```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-fourteen
#| fig-cap: "Probability tree for likelihood of reducing waste, predicted by age and education."

@@

#### Likelihood of sorting waste correctly
fig-fifteen identifies education as the primary factor for sorting waste. Those with higher education (at least community/trade school or postgraduate) are more likely to sort waste correctly, with an error rate of 37.3%. Age further refines the predictions for those with lower education levels, with younger individuals under 37 years being more likely to sort waste correctly, albeit with a higher error rate of 57.1%. Older individuals (37 years and above) are more likely to sort waste correctly, with an error rate of 35.3%. The model emphasizes the role of education in fostering proper waste-sorting behaviors, with age providing additional nuance.
```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-fifteen
#| fig-cap: "Probability tree for likelihood of sorting waste correctly, predicted by age and education."

@@

### Home and Lifestyle Changes
#### Likelihood of making home improvements
fig-sixteen shows that age is the key predictor for home improvement behavior. Individuals aged 36 years and older are predominantly classified as "Already doing this or have done this," with an error rate of 59.6%. For younger individuals (under 36), education plays a role: those with higher educational attainment ("Community/Trade School," "Some University," or "Postgraduate") are "Very likely" to make home improvements, with an error rate of 55%. Conversely, individuals with lower education levels are classified as "Somewhat likely," with a higher error rate of 67.4%. This model highlights how age and education interact to influence home improvement decisions.

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-sixteen
#| fig-cap: "Probability tree for likelihood of making home improvements predicted by age and education."

@@

#### Likelihood of adopting meat alternatives
fig-seventeen demonstrates that age is the primary factor for adopting meat alternatives. Younger individuals (under 35 years) are more likely to adopt meat alternatives, with an error rate of 44%. For individuals aged 35 years or older, education becomes a key predictor. Those with community/trade school, undergraduate, or postgraduate education are more likely to adopt meat alternatives, with varying error rates, including "Somewhat unlikely" for younger individuals at 62.9%. This model highlights how age and education together shape dietary choices.

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-seventeen
#| fig-cap: "Probability tree for likelihood of adopting meat alternatives, predicted by age and education."

@@

#### Likelihood of reducing hydro usage
fig-eighteen reveals that age is the primary predictor influencing hydro usage reduction. Younger individuals (under 35 years) are "Very likely" to reduce hydro usage, with an error rate of 60.9%. Older individuals (35 years and above) are further split by education and age, with those over 64 years also "Very likely" to engage in hydro conservation, with a lower error rate of 37.5%. Middle-aged individuals (35–64 years) with higher education are either "Already doing this or have done this" (error rate 56.3%) or "Somewhat likely" (error rate 50.0%). This model underscores the combined role of age and education in promoting energy-saving behaviors.

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-eighteen
#| fig-cap: "Probability tree for likelihood of reducing hydro usage, predicted by age and education."

@@

@@

## Model Summary Table
fig-nineteen presents the accuracy of models predicting climate-friendly actions based on age and education. A threshold emerges from the data: models with accuracy above 50% are considered reliable, while those below require careful interpretation. Sorting waste correctly, with an accuracy of 59%, is the most reliable prediction, indicating a stronger association between demographic factors and this behavior. Models like home improvements (33%) offer limited predictive value but remain somewhat informative. In contrast, actions such as adopting meat alternatives (25%) or choosing electric or hybrid vehicles (25%) approach random chance, making their predictions less reliable for drawing meaningful conclusions.

```{r}
#| warning: false
#| echo: false
#| message: false
#| results: asis
##| label: fig-nineteen
#| fig-cap: "This table presents the accuracy of models predicting the likelihood of various climate-friendly actions based on age and education."

@@

## Likelihood taking action by age
fig-twenty shows the relationship between age and the likelihood of engaging in various climate-related actions. The x-axis represents the age groups, and the y-axis indicates the percentage of respondents reporting different likelihood levels, from "Already doing this" to "Very unlikely".

@@

```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-twenty
#| fig-cap: "Likelihood of Taking Climate Change Actions by Age Group."

@@

## Likelihood taking action by education 
fig-twenty-one displays how the likelihood of engaging in climate-friendly actions varies by educational attainment. The x-axis represents the proportion of respondents in each education category, while the y-axis lists the educational levels, from low to high educational attainment. Each action is shown across separate facets, revealing how education level correlates with engagement in each behavior.

@@

### Home and Lifestyle Changes
For making home improvements to increase energy efficiency, individuals in the 35-54 age range are the most likely to report having already made improvements or being very likely to do so. In general, at least half of the sample across all age groups reports that they are already engaged in or very likely to make energy-efficient home improvements, although younger individuals (15-24) are less likely to take this action. When it comes to adopting meat alternatives, individuals aged 35 and above are more likely to reduce meat consumption with plant-based alternatives, while those aged 15-24 are the least likely to engage in this behavior. For reducing hydro usage, older age groups, particularly those aged 65+, report the highest rates of already engaging in or being very likely to engage in this action. However, the 65+ age group also reports the highest percentage of being "very unlikely" to reduce hydro usage.
```{r}
#| warning: false
#| echo: false
#| message: false
##| label: fig-twenty-one
#| fig-cap: "Likelihood of Taking Climate Change Actions by Education Level."

@@

>>>>

### Education
#### Individual 2018
Figure 2 shows the distribution of respondents by their highest level of education attained in 2018. The most common education levels are "Completed undergraduate degree" and "Postgraduate/professional school," indicating a well-educated respondent population. In contrast, fewer respondents reported having attended or completed community college, vocational, or trade school, suggesting lower representation in these categories. A small portion of respondents chose not to disclose their education level. Overall, the distribution reflects a skew toward higher levels of educational attainment.
```{r}
#| warning: false
#| echo: false
#| message: false

@@

Figure 3: The histogram shows the distribution of self-reported climate change knowledge among 2018 survey respondents. This distribution provides insights into public awareness and the extent to which individuals feel informed about climate change.

#### Comparison 2018 to 2021
Table 3 illustrates the self-reported level of extent individuals feel informed about climate change and climate action for 2018 and 2021. In 2018, the large majority of individuals say they are very informed with 60 percent and the following category is not very informed with 27 percent. In 2021, The majority of individuals also say they are very informed but to a lesser degree with 48 percent and this is closely followed by only a little informed with 37 percent. There is a trend that individuals in 2018 reported greater levels of knowledge about climate change whereas 2021 illustrates a more even split between very informed and only little informed, portraying that fewer individuals have adequate knowledge about climate change.

Table 3: This table highlights the extent feel informed distribution of survey respondents in 2018 and 2021.

@@

# Print the table (using kable for styling)
informed_summary_combined %>%
  kbl() %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

#### Comparison 2018 to 2021
Table 4 compares the preferred methods of delivering climate change information between the 2018 and 2021 surveys. While the top three methods namely advertising campaigns, newsletters, and the Toronto.ca website remained consistent across both years, there were notable shifts in preferences. Traditional methods such as newsletters, advertising campaigns, brochures, and pamphlets saw a decline in 2021. In contrast, social media platforms including Twitter, Facebook, and Instagram gained traction, reflecting a growing reliance on digital communication channels.

@@

@@

# Print the table (using kable for styling)
communication_summary_combined %>%
  kbl(caption = "This table illustrates the prefered method of communication of survey respondents in 2018 and 2021") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

@@

### Likelihood Taking Action 
Figure 5 illustrates the self-reported likelihood of individuals taking various actions to minimize environmental impacts. The x-axis displays nine different climate actions, while the y-axis represents the percentage of individuals corresponding to each likelihood level. A stacked bar chart is used to visualize the distribution of responses, with a color-coded legend ranging from green (high likelihood of action) to red (low likelihood of action).

@@

@@

#### Likelihood of switching to an electric or hybrid vehicle
Figure 6 reveals that education is the primary predictor for the likelihood of switching to an electric or hybrid vehicle. Individuals with "Some Community/Trade School" or "Some University" education are "Very likely" to invest, though with moderate variability (error rate 63.8%). Those with higher education show more age-based segmentation. Younger individuals under 35 years are more likely to invest, with a lower error rate (30%). Older individuals, particularly those over 69 years, are "Very unlikely" to invest. This model demonstrates how age and education interact to influence vehicle choice.

```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

Figure 6: Probability tree for likelihood of investing in an electric or hybrid vehicle, predicted by age and education

#### Likelihood of minimizing car use
Figure `7` shows that the likelihood of minimizing car use is consistent across the sample, with a high error rate (61.9%). The model classifies all individuals into the "Already doing this or have done this" category, indicating widespread adoption of this behavior, regardless of age or education

```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

Figure 7: Probability tree for likelihood of minimizing car use, predicted by age and education.

#### Likelihood of walking or cycling shorter distances
Figure `8` indicates that education is the primary factor influencing short-distance walking or cycling. Those with higher education (postgraduate, undergraduate, or some university) are more likely to engage in this behavior, with an error rate of 51.9%. For those with lower education, age plays a significant role, with younger individuals (under 66 years) more likely to walk or cycle, with an error rate of 60%. This model highlights how education and age together shape short-distance mobility, with error rates reflecting the model's predictive variability.

```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

```
Figure 8: Probability tree for likelihood of walking or cycling shorter distances, predicted by age and education.

### Waste and Product Actions
#### Likelihood of purchasing green products
Figure `9` reveals that age is the initial split for green product purchasing behavior. Younger individuals (under 38 years) are more likely to purchase green products, with an error rate of 51.9%. Education further refines the prediction. Older individuals (over 38 years) are classified by both age and education, with postgraduate education leading to higher engagement in purchasing green products, with an error rate of 60%. This model shows how age and education influence consumer choices in sustainable products.
```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

FIgure 9: Probability tree for likelihood of purchasing green products, predicted by age and education

#### Likelihood of reducing waste by repurposing used product
Figure `10` shows that age is the primary factor for waste reduction behaviors. Younger individuals (under 35 years) are more likely to repurpose waste, with an error rate of 57.1%. Education further refines the classification. Older individuals (over 35 years) show varied likelihoods based on education, with more highly educated individuals more likely to repurpose waste, with an error rate of 60%. This highlights the influence of both age and education on waste-reduction habits.
```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

Figure 10: Probability tree for likelihood of reducing waste, predicted by age and education.

#### Likelihood of sorting waste correctly
Figure `11` identifies education as the primary factor for sorting waste. Those with higher education (at least community/trade school or postgraduate) are more likely to sort waste correctly, with an error rate of 37.3%. Age further refines the predictions for those with lower education levels, with younger individuals under 37 years being more likely to sort waste correctly, albeit with a higher error rate of 57.1%. Older individuals (37 years and above) are more likely to sort waste correctly, with an error rate of 35.3%. The model emphasizes the role of education in fostering proper waste-sorting behaviors, with age providing additional nuance.
```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

Figure 11: Probability tree for likelihood of sorting waste correctly, predicted by age and education

### Home and Lifestyle Changes
#### Likelihood of making home improvements
Figure `12` shows that age is the key predictor for home improvement behavior. Individuals aged 36 years and older are predominantly classified as "Already doing this or have done this," with an error rate of 59.6%. For younger individuals (under 36), education plays a role: those with higher educational attainment ("Community/Trade School," "Some University," or "Postgraduate") are "Very likely" to make home improvements, with an error rate of 55%. Conversely, individuals with lower education levels are classified as "Somewhat likely," with a higher error rate of 67.4%. This model highlights how age and education interact to influence home improvement decisions.

```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

Figure 12: Probability tree for likelihood of making home improvements predicted by age and education.

#### Likelihood of adopting meat alternatives
Figure `13` demonstrates that age is the primary factor for adopting meat alternatives. Younger individuals (under 35 years) are more likely to adopt meat alternatives, with an error rate of 44%. For individuals aged 35 years or older, education becomes a key predictor. Those with community/trade school, undergraduate, or postgraduate education are more likely to adopt meat alternatives, with varying error rates, including "Somewhat unlikely" for younger individuals at 62.9%. This model highlights how age and education together shape dietary choices.

```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

Figure 13: Probability tree for likelihood of adopting meat alternatives, predicted by age and education.

#### Likelihood of reducing hydro usage
Figure `14` reveals that age is the primary predictor influencing hydro usage reduction. Younger individuals (under 35 years) are "Very likely" to reduce hydro usage, with an error rate of 60.9%. Older individuals (35 years and above) are further split by education and age, with those over 64 years also "Very likely" to engage in hydro conservation, with a lower error rate of 37.5%. Middle-aged individuals (35–64 years) with higher education are either "Already doing this or have done this" (error rate 56.3%) or "Somewhat likely" (error rate 50.0%). This model underscores the combined role of age and education in promoting energy-saving behaviors.

```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "100%"
#| out.height: "auto"  # Maintain aspect ratio

@@

Figure 14: Probability tree for likelihood of reducing hydro usage, predicted by age and education.

@@

## Model Summary Table
Table `5` presents the accuracy of models predicting climate-friendly actions based on age and education. A threshold emerges from the data: models with accuracy above 50% are considered reliable, while those below require careful interpretation. Sorting waste correctly, with an accuracy of 59%, is the most reliable prediction, indicating a stronger association between demographic factors and this behavior. Models like home improvements (33%) offer limited predictive value but remain somewhat informative. In contrast, actions such as adopting meat alternatives (25%) or choosing electric or hybrid vehicles (25%) approach random chance, making their predictions less reliable for drawing meaningful conclusions.

```{r}
#| warning: false
#| echo: false
#| message: false
#| results: asis
#| fig-cap: "Table 5: This table presents the accuracy of models predicting the likelihood of various climate-friendly actions based on age and education."

@@

## Likelihood taking action by age
Figure `15` shows the relationship between age and the likelihood of engaging in various climate-related actions. The x-axis represents the age groups, and the y-axis indicates the percentage of respondents reporting different likelihood levels, from "Already doing this" to "Very unlikely".

@@

```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "200%"
#| out.height: "auto"

@@

Figure 15: The facet of stacked bar charts allows for comparison of which climate-friendly actions are most practiced amongts different age categories.

## Likelihood taking action by education 
Figure `16` displays how the likelihood of engaging in climate-friendly actions varies by educational attainment. The x-axis represents the proportion of respondents in each education category, while the y-axis lists the educational levels, from low to high educational attainment. Each action is shown across separate facets, revealing how education level correlates with engagement in each behavior.

@@

### Home and Lifestyle Changes
For making home improvements to increase energy efficiency, individuals in the 35-54 age range are the most likely to report having already made improvements or being very likely to do so. In general, at least half of the sample across all age groups reports that they are already engaged in or very likely to make energy-efficient home improvements, although younger individuals (15-24) are less likely to take this action. When it comes to adopting meat alternatives, individuals aged 35 and above are more likely to reduce meat consumption with plant-based alternatives, while those aged 15-24 are the least likely to engage in this behavior. For reducing hydro usage, older age groups, particularly those aged 65+, report the highest rates of already engaging in or being very likely to engage in this action. However, the 65+ age group also reports the highest percentage of being "very unlikely" to reduce hydro usage.
```{r}
#| warning: false
#| echo: false
#| message: false
#| out.width: "200%"
#| out.height: "auto"

@@

Figure 16: The facet of stacked bar charts allows for comparison of which climate-friendly actions are most practiced amongts different education attainments.
